---
title: "ggplot_2"
author: "Clement Mugenzi"
date: "9/26/2019"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
```

```{r cache=TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
view(weather_df)
```


## Create ggplot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

alternative way of making this plot

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()
```


saving initial plot: mostly i dont use this

```{r}
scatterplot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

scatterplot
```

adding color

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .4) # alpha is for transparancy
```



why do `aes` positions matter?

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .4) +
  geom_smooth(se = F) # se makes the plot cleaner
```


facets

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .4) +
  geom_smooth(se = F) +
  facet_grid(~name)
```



this is fine. but not interesting

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(size = 2, se = F)
```


## Some extra stuff


```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_smooth(size = 2, se = F)
```


2d density

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_bin2d()
```

## More kinds of plots

```{r}
# Fill makes sure the bars are filled with colors
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram(position = "dodge", binwidth = 2) +
  facet_grid(~name)
```

Density plots!!

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = .3)
```


```{r}
# univariate plots
weather_df %>% 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```


```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmax, fill = name)) +
  geom_violin()
```

ridge plots!!

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges()
```


## saving Plots

```{r}
ggp_ridge_temp =
weather_df %>% 
  ggplot(aes(x = tmax, y = name)) +
  geom_density_ridges()

ggsave("ggp_ridge_temp.pdf",
       ggp_ridge_temp)
```


```{r, fig.width=6}
 weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point()

```




```{r, fig.width=12}
 weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point()

```


## Making New plots

start with an old plot

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5)
```

Making labels


```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  )
```


x axis ticks marks (to highlight where you wanna focus on your graph)


```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) +
  scale_x_continuous(
    breaks = c(-15, -5, 20),
    labels = c("-15C", "-5(still too cold)", "20C") # you can do whatever
  ) +
  scale_y_continuous(
    trans = "sqrt"
  )
```


## Change the default colors

```{r}
ggp_base = weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) +
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )
ggp_base
```

## themes

```{r}
ggp_base +
  theme_minimal() +
  theme(legend.position = "bottom") 
# to make some space and here order matters a lot
```

## More than one dataset

```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(data = waikiki, aes(x = date, y = tmax,
                           color = name)) +
  geom_point(aes(size = prcp)) +
  geom_line(data = central_park)
```


(Brief aside about colors)

```{r}
waikiki %>% 
  ggplot(aes(x = date, y = tmax)) +
  geom_point(color = "blue") # here it matters where color = is placed.
```


## Multi-panel plots using patchwork (putting multiple plots all in the same graph)

```{r}
tmax_tmin_p = 
  weather_df %>% 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df %>% 
  filter(prcp > 0) %>% 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = 
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```


## Data Manipulation

```{r}
weather_df %>% 
  mutate(
    name = factor(name),
    name = fct_reorder(name, tmax) #reorder the factors
  ) %>% 
  ggplot(aes(x = name, y = tmax, color = name)) +
  geom_boxplot()
```


## Restructure then plot

```{r}
weather_df %>% 
  pivot_longer(
    tmax:tmin,
    names_to = "observation",
    values_to = "temperature"
  ) %>% 
  ggplot(aes(x = temperature, fill = observation)) +
  geom_density() +
  facet_grid(~name) +
  theme(legend.position = "bottom")
```


litters and pups

```{r}
pup_data = 
  read_csv("Data/data_import_examples/FAS_pups.csv", col_types = "ciiiii") %>%
  janitor::clean_names() %>%
  mutate(sex = recode(sex, `1` = "male", `2` = "female")) 

litter_data = 
  read_csv("Data/data_import_examples/FAS_litters.csv", col_types = "ccddiiii") %>%
  janitor::clean_names() %>%
  select(-pups_survive) %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  mutate(wt_gain = gd18_weight - gd0_weight,
         day_of_tx = as.numeric(day_of_tx))

fas_data = left_join(pup_data, litter_data, by = "litter_number") 

fas_data %>% 
  select(sex, dose, day_of_tx, pd_ears:pd_walk) %>% 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome", 
    values_to = "pn_day") %>% 
  drop_na() %>% 
  mutate(outcome = forcats::fct_reorder(outcome, day_of_tx, median)) %>% 
  ggplot(aes(x = dose, y = pn_day)) + 
  geom_violin() + 
  facet_grid(day_of_tx ~ outcome)
```






















